A weakly structured stem for human origins in Africa

### 2022-02-24 paper.tex
cleaned up fig references and relate paragraph

We used \texttt{Relate} \citep{Speidel2019-nj} to infer the distribution of
coalescence rates over time, which are commonly interpreted as effective
population size histories. Trajectories of $N_e$ from prior studies strongly
support a bottleneck in Out of Africa populations. African populations also dip
in effective size around 100-200kya, although this decrease is more modest than
the OOA event \citep{Mallick2016-lx}. One possible explanation is that
ancestral population structure during the Middle Stone Age inflated $N_e$ and
subsequent gene flow among ancestral Africans resulted in a more recent
decreased $N_e$ \citep{Mazet2016-wn}.
%To our knowledge, explicit comparisons of this hypothesis with SMC-style
%approaches have not been published.
We simulated genomic data under four inferred demographic models to compare
with observed coalescence rate histories. All models, including the
single-origin model, generate multi-modal curves recapitulating an inferred ancestral
increase in $N_e$ between 100kya-1Mya. Thus the reduction in inverse coalescence rate 
is captured in our model, but is primarily explained by population structure in the best-fitting models \sgcomment{is that fair?}.     
None of the models perfectly match the exact timing of
the early fluctuations in $N_e$ observed in the data \ref{fig:supp-iicr-data-subset}. 
%Simulated Relate Ne curves under the CM and MM are distinguished by having 2
%and 3 modes, respectively.  The Continuous Migration model had the better
%qualitative fit with the data, but the ancestral mode was shifted further back
%in time relative to the data.
Each simulated model also underestimated the level of population growth during
the Holocene, indicating either artifacts of data inference in real data or that additional free parameters are necessary to refine
these growth rates. \sgcomment{I am not sure I am happy with these interpretations 
-- let's discuss. I feel like this paragraph does not support our data well, and a likely culprit 
in explaining the differences is the inaccuracy of TMRCA estimates from real data.
Also, we would need a link to a supplementary figure!}

>>>>

We used \texttt{Relate} \citep{Speidel2019-nj} to infer the distribution of
coalescence rates over time, which are commonly interpreted as effective
population size histories. Trajectories of $N_e$ from prior studies strongly
support a bottleneck in Out of Africa populations. African populations also dip
in effective size around 100-200kya, although this decrease is more modest than
the OOA event \citep{Mallick2016-lx}. One possible explanation is that
ancestral population structure during the Middle Stone Age inflated $N_e$ and
subsequent gene flow among ancestral Africans resulted in a more recent
decreased $N_e$ \citep{Mazet2016-wn}.
%To our knowledge, explicit comparisons of this hypothesis with SMC-style
%approaches have not been published.
We simulated genomic data under four inferred demographic models to compare
with observed coalescence rate histories. All models, including the
single-origin model, generate multi-modal curves recapitulating an inferred ancestral
increase in $N_e$ between 100kya-1Mya (Figure \ref{fig:supp-iicr-sim}). Thus the reduction in inverse coalescence rate 
is captured in our model, but is primarily explained by population structure in the best-fitting models \sgcomment{is that fair?}.     
None of the models perfectly match the exact timing of
the early fluctuations in $N_e$ observed in the data (Figure \ref{fig:supp-iicr-data-subset}). 
\sgcomment{Perhaps: These discrepancies are comparable to discrepancies observed in relate curves inferred 
for the same populations using different subsets of the data (see Figures \ref{ig:supp-iicr-data-subset} and \ref{ig:supp-iicr-data-subset})}
%Simulated Relate Ne curves under the CM and MM are distinguished by having 2
%and 3 modes, respectively.  The Continuous Migration model had the better
%qualitative fit with the data, but the ancestral mode was shifted further back
%in time relative to the data.
Each simulated model also underestimated the level of population growth during
the Holocene, indicating either artifacts of data inference in real data or that additional free parameters are necessary to refine
these growth rates. \sgcomment{Perhaps: There are other discrepancies between odel }

\sgcomment{I am not sure I am happy with these interpretations 
-- let's discuss. I feel like this paragraph does not support our data well, and a likely culprit 
in explaining the differences is the inaccuracy of TMRCA estimates from real data.
Also, we would need a link to a supplementary figure!}
